Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-11-04 qcd.tex
first proofread

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experiment.
    From (\ref{eqn:dsigma}), the partonic cross-section
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$, (\ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

    In this chapter we will discuss the problems that
    arise when evaluating these Feynman diagrams and outline
    some solutions that have been widely adopted to circumvent
    these issues to provide real-world applicable predictions.
    This chapter forms the theoretical foundations on which
    the work of this thesis is built upon.

@@

\section{Divergent structures}\label{sec:divergences}
    We have seen that matrix element computations
    result in evaluating Feynman diagrams. From
    the QCD Feynman rules (ref here), it becomes
    clear that integrating over all momentum modes
    will give rise to singularities. The divergences
    associated with high energy modes are ultraviolet
    (UV) divergences. On the opposite end of the energy
    spectrum, low energy modes also give rise to 
    infrared (IR) divergences.

@@

\subsection{Ultraviolet divergences}\label{sec:UV}
    During intermediate steps of calculations,
    such as trying to compute loop diagrams seen in
    the Figure~\ref{fig:bubble_diagram},
    we have to evaluate integrals of the form
    \begin{equation}\label{eqn:loop_integral}
        I_{\mathrm{UV}} = \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
    \end{equation}
    where a cutoff scale $\Lambda$ has been introduced to
    capture the divergence as the loop momenta
    $k \rightarrow \infty$. It is clear that the
    integral diverges in this high energy limit,
    hence the name ultraviolet divergence.
    This UV divergence cannot influence the
    experimentally observable quantities at experiments,
    which are finite, because the final observable
    should not depend on the intermediate steps taken
    to get there.
    Infinities associated with UV divergences
    are removed via the process of renormalisation
    which introduces counterterms in the Lagrangian
    to exactly cancel these divergences.

    \begin{figure}
        \includegraphics[scale=1.2]{qcd/bubble_diagram.pdf}
        \caption{Bubble diagram}
        \label{fig:bubble_diagram}
    \end{figure}

@@

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        % \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}(-p^{2})^{-\epsilon}}{(4\pi)^{2-\epsilon}}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)}\Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

\subsection{Renormalisation}\label{sec:renormalisation}
    With the UV divergence regularised by dimensional
    regularisation, it is now possible to construct the forms
    of the counterterms order-by-order in $\alpha_{s}$ to
    explicitly cancel them.

    A subtlety with adjusting the dimension of the theory
    from $d=4$ to arbitrary dimensions is that the mass dimension
    of the Lagrangian has to change accordingly to retain a
    dimensionless action. We can account for this by
    introducing an arbitrary scale $\mu_{R}$, the renormalisation
    scale, into the gauge coupling
    constant to modify the mass dimension of the Lagrangian.
    For QCD this is done via the modification
    \begin{equation}\label{eqn:dimensionful_coupling}
        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
    \end{equation}
    where it is understood the original gauge coupling is
    dimensionless. The order of $\mu_{R}$ is determined by examining
    the dimensions of the gauge and fermion fields
    in the Lagrangian (\ref{eqn:L_QCD}).
    We see that this introduces a scale dependence
    on the coupling constant $\alpha_{s}$, a point we will return
    to in \ref{sec:alpha_running}.
    % In a calculation,
    % intermediate steps will have a dependence on $\mu$,
    % however, once the $\epsilon \rightarrow 0$ limit
    % is taken the scale dependence drops out.

    The inclusion of counterterms into the theory can be
    thought of as renormalisations of the the bare fields
    and couplings in (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see e.g. Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{\mathrm{renorm}} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals (\ref{eqn:gamma_laurent}). However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

    In an all orders calculation of an observable,
    there would be no dependence on the renormalisation
    scale as it is a remnant of the regularisation prescription
    However, because in perturbative QCD observables
    are calculated order-by-order, there will be a residual
    dependence on the renormalisation scale stemming from
    the missing higher order terms. The conventional way
    that the community estimates the uncertainty associated
    with this residual dependence is by carrying out
    a scale variation. This involves evaluating the
    relevant quantities at the nominal renormalisation
    scale and again with the scale varied by a factor of
    two above and below.

\subsection{Running of the coupling constant}\label{sec:alpha_running}
    The bare coupling should not depend on the actual
    value of the renormalisation scale,
    \begin{equation}\label{eqn:d_g_bare_d_mu}
        \dfrac{\mathrm{d}g_{\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
    \end{equation}
    or said another way, the renormalisation process
    is independent of the renormalisation scale. The
    consequence of this is that $\alpha_{s}$ has to
    depends on the renormalisation scale.

    The behaviour of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu^{2} \dfrac{\partial \alpha_{s}(\mu^{2})}{\partial \mu^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive,
    which is what we have observed in nature with six
    flavours of quarks. This behaviour of positive
    $\beta_{n}$ is also the case in all coefficients
    computed to date, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

@@

    Another consequence of the running coupling
    is colour confinement, in which quarks and
    gluons cannot exist as free particles. Due
    to the increasing strength of coupling at low
    energies, quarks and gluons are forced to
    form composite, colourless particles. This means
    that at low energies, perturbative QCD
    is not an accurate description of nature,
    and the behaviour of hadrons cannot be predicted
    from perturbation theory.

\subsection{Infrared divergences}\label{sec:IR_divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    in loop integrals (virtual) and when an emission
    of an extra particle (c.f. splitting functions
    (\ref{eqn:AP_kernels})) has vanishing energy (soft),
    or becomes parallel to an external colourful line (collinear).

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{v} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{v} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
    corresponding to the soft and collinear divergence.

    \begin{figure}
        \includegraphics[scale=1.2]{qcd/vertex_correction.pdf}
        \caption{Vertex correction}
        \label{fig:vertex_correction}
    \end{figure}

@@

    In DR, the singularities from soft and collinear divergences
    are manifest as $\epsilon$ poles which makes the cancellation
    simple between the real and virtual parts. However, in practice, 
    the phase-space integrals of the matrix elements are rarely carried
    out analytically due to the high dimensionality of the integral.
    Instead they are done numerically with the use
    of Monte Carlo methods, which requires the integral to be in
    integer dimensions. Therefore, numerical techniques
    have been developed to deal with these IR divergences as well. The
    most common technique, subtraction, will be discussed in
    more detail in Section~\ref{sec:subtraction}.

@@

    The regions of phase-space in which real emission
    matrix elements diverge are when the emission is
    soft and/or collinear. It can be shown that in these
    regions of phase-space, the matrix element factorises
    into a process-independent singular factor, multiplied
    by a matrix element with the unresolved parton
    being absorbed by an emitting leg (reduced matrix
    element). This factorisation, however, is not exact
    in QCD where there are spin and colour correlations.

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \bm{S}_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $\bm{S}_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $\bm{S}_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the $n+1$ external states.

@@

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element is obtained by removing
    the soft gluon. The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors to give factors of
    $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

@@

    Similar to (\ref{eqn:soft_factorisation}), (\ref{eqn:F_collinear_factorisation})
     and (\ref{eqn:I_collinear_factorisation}) are only
    true in the strict collinear limit. Away from these
    limits care has to be taken to conserve momenta
    via the use of momenta mappings.

>>>>

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Quantum chromodynamics in practice}
\label{chapter:qcd}
    In the previous chapter we introduced the basic concepts
    of QCD and setup the framework of calculating
    cross-sections. Cross-sections are the most relevant
    quantities to compute as they can be measured
    experimentally, and are a property of the particles
    being collided, rather than depending on the
    specifics of the experimental procedure.
    This provides a bridge to compare
    theoretical predictions and experimental measurements.
    The relevant quantity calculated in perturbation theory
    is the partonic cross-section which
    is the matrix element integrated over the relevant
    final-state phase-space, normalised by a flux factor.
    The matrix elements are
    calculated order-by-order in the strong coupling
    $\alpha_{s}$ (c.f. \ref{eqn:matrix_element}),
    where each order corresponds to a set of Feynman diagrams
    that have the appropriate number of vertices,
    as each QCD vertex brings along a factor of $\alpha_{s}$.

    In this chapter we will discuss the problems that
    arise when evaluating these Feynman diagrams and outline
    some solutions that have been widely adopted to circumvent
    these issues in order to provide real-world applicable predictions.
    This chapter forms the theoretical foundations upon which
    this thesis is built on.

@@

\section{Divergent structures}\label{sec:divergences}
    The computation of matrix elements essentially reduces down
    to evaluating Feynman diagrams which are analytical expressions
    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
    \ref{fig:qcd_feynman_rules}). In evaluating certain
    topologies of Feynman diagrams, integral expressions
    containing unconstrained momenta will give rise to
    singularities. The divergences
    associated with high energy modes are ultraviolet
    (UV) divergences and on the opposite end of the energy
    spectrum, low energy modes give rise to 
    infrared (IR) divergences.

@@

\subsection{Ultraviolet divergences}\label{sec:UV}
    During intermediate steps of calculations,
    such as the compututation of loop diagrams seen in
    Figure~\ref{fig:bubble_diagram},
    we have to evaluate integrals of the form
    \begin{equation}\label{eqn:loop_integral}
        I_{\mathrm{UV}} = \int_{0}^{\Lambda} \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} \sim \log{\Lambda} \, ,
    \end{equation}
    where a cutoff scale $\Lambda$ has been introduced to
    capture the divergence as the loop momenta
    $\ell \rightarrow \infty$. It is clear that the
    integral diverges in this high energy limit,
    hence the name ultraviolet divergence.
    This UV divergence cannot influence the
    experimentally observable quantities at experiments,
    which are finite, because the final observable
    should not depend on the intermediate steps taken
    to get there.
    Infinities associated with UV divergences
    are removed via the process of renormalisation
    which introduces counterterms in the Lagrangian
    to exactly cancel these divergences.

    \begin{figure}
        \includegraphics[scale=1.2]{qcd/bubble_diagram.pdf}
        \caption{Massless bubble diagram which is an example of where unconstrained
        loop momenta can lead to UV divergences.}
        \label{fig:bubble_diagram}
    \end{figure}

@@

    In DR we define $d = 4 - 2\epsilon$ such that
    after Feynman parametrisation (\ref{eqn:loop_integral})
    becomes
    \begin{equation}\label{eqn:dim_reg}
        I_{\mathrm{UV}} = \int \dfrac{\mathrm{d}^{d}\ell}{(2\pi)^{d}}\dfrac{1}{\ell^{2}(\ell+p)^{2}} = \dfrac{\mathrm{i}}{(4\pi)^{2}}\left(\dfrac{-p^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1-\epsilon)^{2}}{\Gamma(2-2\epsilon)} \, \Gamma(\epsilon) \, ,
    \end{equation}
    where the divergence for $d = 4$, or equivalently
    $\epsilon \rightarrow 0$, is now captured in the
    Gamma function, $\Gamma(\epsilon)$. This becomes clear once
    we expand $\Gamma(\epsilon)$ around small $\epsilon$
    \begin{equation}\label{eqn:gamma_laurent}
        \Gamma(\epsilon) = \dfrac{1}{\epsilon} - \gamma_{E} + \dfrac{1}{2}\left(\gamma_{E}^{2} + \dfrac{\pi^{2}}{6}\right)\epsilon + \mathcal{O}(\epsilon^{2}) \, ,
    \end{equation}
    where $\gamma_{E} \approx 0.577$ is the Euler-Mascheroni
    constant, meaning the divergence is now regularised as a pole in
    $\epsilon$.

\subsection{Renormalisation}\label{sec:renormalisation}
    With the UV divergence regularised by dimensional
    regularisation, it is now possible to construct the
    counterterms order-by-order in $\alpha_{s}$ to
    explicitly cancel the divergences.

    A subtlety with adjusting the dimension of the theory
    from $d=4$ to arbitrary dimensions is that the mass dimension
    of the Lagrangian has to change accordingly to retain a
    dimensionless action in natural units. We can account for this by
    introducing an arbitrary energy scale $\mu_{R}$, the renormalisation
    scale, into the gauge coupling
    constant to modify the mass dimension of the Lagrangian.
    For QCD this is done via the modification
    \begin{equation}\label{eqn:dimensionful_coupling}
        g_{s} \rightarrow g_{s}\mu_{R}^{\epsilon} \, .
    \end{equation}
    where it is understood that the original gauge coupling is
    dimensionless. The order of $\mu_{R}$ is determined by examining
    the dimensions of the gauge and fermion fields
    in the Lagrangian (\ref{eqn:L_QCD}).
    We see that this introduces a scale dependence
    on the coupling constant $\alpha_{s}$, a point we will return
    to in Section~\ref{sec:alpha_running}.
    % In a calculation,
    % intermediate steps will have a dependence on $\mu$,
    % however, once the $\epsilon \rightarrow 0$ limit
    % is taken the scale dependence drops out.

    The inclusion of counterterms into the theory can be
    thought of as renormalisations of the bare fields
    and couplings in the QCD Lagrangian (\ref{eqn:L_QCD}) to restore predictive
    power to the theory. The renormalised fields and couplings
    can be given as (see for instance Ref.~\cite{mangano1999introduction})
    \begin{align}\label{eqn:renormlised_fields}
        \psi_{\mathrm{bare}} &= \sqrt{Z_{2}} \psi_{R} \, , \nonumber \\
        A_{\mathrm{bare}}^{\mu} &= \sqrt{Z_{3}} A_{R}^{\mu} \, ,\\
        g_{s,\mathrm{bare}} &= Z_{g} \mu_{R}^{\epsilon} g_{s,R} \, , \nonumber
    \end{align}
    where it conventional to define $Z_{1} = Z_{g}Z_{2}\sqrt{Z_{3}}$
    such that we can set $Z_{n} = 1 + \delta_{n}$ for $n \in \{1, 2, 3\}$.
    In this way we can write the Lagrangian as
    \begin{equation}
        \mathcal{L}_{R} = \mathcal{L}_{\mathrm{bare}} + \mathcal{L}_{\mathrm{c.t.}} \, ,
    \end{equation}
    where the counterterms appearing in $\mathcal{L}_{\mathrm{c.t.}}$
    are determined by calculating $\delta_{n}$. There is freedom
    in the choice of $\delta_{n}$ which gives rise to
    different regularisation schemes. The minimal subtraction
    (MS) scheme subtracts only the epsilon pole appearing
    in loop integrals. However, the most
    commonly used regularisation scheme is the modified minimal
    subtraction ($\overline{\mathrm{MS}}$) scheme which subtracts the
    epsilon pole plus a universal constant appearing in all
    loop integrals.

    In an all orders calculation of an observable,
    there would be no dependence on the renormalisation
    scale as it is a remnant of the regularisation prescription
    However, because in perturbative QCD observables
    are calculated order-by-order, there will be a residual
    dependence on the renormalisation scale stemming from
    the missing higher order terms. The conventional way
    that the community estimates the uncertainty associated
    with this residual dependence is by carrying out
    a scale variation (see Section~\ref{sec:sec:scale_variations}).

\subsection{Running of the coupling constant}\label{sec:alpha_running}
    The renormalisation scale introduced during regularisation
    is a mathematical artifact and should not impact any measurable
    quantities. Therefore, the bare coupling should not depend on
    the renormalisation scale,
    \begin{equation}\label{eqn:d_g_bare_d_mu}
        \dfrac{\mathrm{d}g_{s,\mathrm{bare}}}{\mathrm{d}\mu_{R}} = 0 \, ,
    \end{equation}
    or said another way, the renormalisation process
    is independent of the actual value of the renormalisation scale.
    The consequence of this is that the renormalised coupling, $\alpha_{s}(\mu_{R})$, has to
    depend on the renormalisation scale instead.

    The renormalisation scale dependence of $\alpha_{s}$
    is governed by the Callan-Symanzik \cite{Callan:1970yg,Symanzik:1970rt}
    $\beta$-function
    \begin{equation}\label{eqn:beta_fn}
        \mu_{R}^{2} \dfrac{\partial \alpha_{s}(\mu_{R}^{2})}{\partial \mu_{R}^{2}} = \beta(\alpha_{s}) \, ,
    \end{equation}
    where the $\beta$-function can be written as a perturbative
    expansion in $\alpha_{s}$
    \begin{equation}\label{eqn:beta_series}
        -\beta(\alpha_{s}) = \alpha_{s}\sum_{n=0}^{\infty}\left(\dfrac{\alpha_{s}}{4\pi}\right)^{n+1} \beta_{n}\, 
    \end{equation}
    where the coefficients $\beta_{n}$ have been computed
    up to $\beta_{4}$ \cite{Baikov:2016tgj,Luthe:2017ttg}.
    The solution to (\ref{eqn:beta_fn}) to first order
    is
    \begin{equation}\label{eqn:1l_alpha}
        \alpha_{s}(\mu_{R}^{2}) = \dfrac{1}{\frac{\beta_{0}}{4\pi}\log\left(\frac{\mu_{R}^{2}}{\Lambda_{\mathrm{QCD}}^{2}}\right)} \, ,
    \end{equation}
    where $\Lambda_{\mathrm{QCD}}$ is the QCD scale,
    the scale at which $\alpha_{s}$ becomes
    large enough that perturbation theory breaks down,
    and $\beta_{0} = (11C_{A} - 4T_{R}n_{f})/3$.
    In QCD, $C_{A}=3$ and $T_{R}=\frac{1}{2}$, meaning for
    $n_{f} < \frac{33}{2}$ the sign of $\beta_{0}$ is positive.
    In nature we have observed six flavours of quarks, so
    $\beta_{0} > 0$. In fact all $\beta$-coefficients
    computed to date are positive, meaning that $\alpha_{s}$
    decreases with increasing energy due to the minus sign
    in (\ref{eqn:beta_series}). This property is known as
    asymptotic freedom \cite{Gross:1973id,Politzer:1973fx},
    and is justification for treating QCD perturbatively
    when the energy scale is high, such as at collider experiments.
    The running of the coupling constant has been
    observed experimentally as illustrated in Figure~\ref{fig:alpha_s_running}.

@@

    Another consequence of the running coupling
    is colour confinement, in which quarks and
    gluons cannot exist as free particles. Due
    to the increasing strength of coupling at low
    energies, quarks and gluons are forced to
    form composite, colourless particles. This means
    that at low energies, perturbative QCD
    is not an accurate description of nature,
    and the behaviour of hadrons cannot be predicted
    from perturbation theory. Instead phenomenological
    models have to be used to describe these processes.

\subsection{Infrared divergences}\label{sec:IR_divergences}
    In Section~\ref{sec:UV} we saw how divergences
    stemming from high energy behaviour would occur
    in loop diagrams. On the other end of the energy
    spectrum we also have divergences from low energy
    modes. Since these divergences emerge from low energy
    behaviour they are called infrared (IR) divergences. There
    are two cases in which these divergences arise:
    \begin{itemize}
        \item virtual divergences in loop integrals,
        \item real emission divergences when an emission
    of an extra particle has vanishing energy (soft),
    or becomes parallel to an external leg (collinear).
    \end{itemize}

    Virtual IR divergences can be understood by looking
    at an integral of the form
    \begin{equation}\label{eqn:virtual_IR_integral}
        I_{\mathrm{V}} = \int \dfrac{\mathrm{d}^{4}\ell}{(2\pi)^{4}} \dfrac{1}{\ell^{2}(\ell+p_{1})^{2}(\ell-p_{2})^{2}} \, ,
    \end{equation}
    which is encountered when calculating the one-loop
    virtual correction to the gluon-quark-antiquark vertex
    as shown in Figure~\ref{fig:vertex_correction}.
    It is clear that the denominator vanishes when $\ell \rightarrow 0$
    or when either $(\ell + p_{1})^{2}$ or $(\ell - p_{2})^{2} \rightarrow 0$.
    These situations correspond to the gluon in the loop propagator
    going soft, or collinear to the external quark or antiquark.
    These divergences are regulated in DR to give
    \begin{equation}\label{eqn:virtual_IR_DR}
        I_{\mathrm{V}} = \dfrac{\mathrm{i}}{(4\pi)^{2}Q^{2}}\left(\dfrac{-Q^{2}}{4\pi}\right)^{-\epsilon}\dfrac{\Gamma(1+\epsilon)\Gamma(1-\epsilon)^{2}}{\Gamma(1-2\epsilon)}\left[\dfrac{1}{\epsilon^{2}}\right] \, ,
    \end{equation}
    where $Q = (p_{1} + p_{2})$. There is a double $\epsilon$ pole manifest,
    corresponding to the soft and collinear divergence.

    \begin{figure}
        \includegraphics[scale=1.2]{qcd/vertex_correction.pdf}
        \caption{One-loop vertex correction for the gluon-quark-antiquark vertex which diverges
        for $\ell \rightarrow 0$.}
        \label{fig:vertex_correction}
    \end{figure}

@@

    In DR, the singularities from soft and collinear divergences
    are manifest as $\epsilon$ poles which makes the cancellation
    simple between the real and virtual parts. However, in practice, 
    the phase-space integrals of the matrix elements are rarely carried
    out analytically due to the high dimensionality of the integral,
    rendering them intractable.
    Instead they are done numerically through the use
    of Monte Carlo methods (see Chapter~\ref{chap:MC}),
    which requires the integral to be in
    integer dimensions. Therefore, numerical techniques
    have been developed to deal with these IR divergences as well. The
    most common technique, subtraction, will be discussed in
    more detail in Section~\ref{sec:subtraction}.

@@

    The regions of phase-space in which real emission
    matrix elements diverge are when the emission is
    soft and/or collinear. It can be shown that in these
    regions of phase-space, the matrix element factorises
    into a process-independent singular factor, multiplied
    by a matrix element with the unresolved parton
    being absorbed by an emitting leg -- we call this the reduced matrix
    element. This factorisation, however, is not exact
    in QCD where there are spin and colour correlations.

    Schematically, the $(n+1)$-body matrix element
    factorises as
    \begin{equation}\label{eqn:me_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow \bm{S}_{ijk} \otimes |\mathcal{M}_{n}|^{2} \, ,
    \end{equation}
    where $\bm{S}_{ijk}$ is a universal singular factor capturing
    the IR divergent behaviour and $|\mathcal{M}_{n}|^{2}$
    is the reduced matrix element. The $\otimes$
    represents the colour and spin correlations existing
    between the singular function and the reduced matrix element.
    This singular factor
    is not unique and can be represented by different
    approximations as long as they reproduce the
    correct IR behaviour. In this thesis we discuss
    two approximations: Catani-Seymour dipoles \cite{Catani:1996vz}
    in Section~\ref{sec:CS_dipoles} and antenna functions \cite{Gehrmann-DeRidder:2005btv}
    in Section~\ref{sec:antenna_functions}. The indices
    of $\bm{S}_{ijk}$ already hints at the fact that the singular
    functions only depends on three partons, and not all
    of the external states.

@@

\subsection{Soft limits}\label{sec:me_soft}
    Consider a tree-level matrix element $|\mathcal{M}_{n+1}|^{2}$
    with a final-state gluon $j$. Note that reduced matrix
    elements associated with taking a quark soft have to vanish
    due to violation of quark number, therefore only gluons
    are considered.
    The limit of the soft gluon with momentum
    $p_{j}$ can be parametrised by
    \begin{equation}\label{eqn:soft_parametrisation}
        p_{j}^{\mu} = \lambda q^{\mu} \, , \quad \lambda \rightarrow 0 \, ,
    \end{equation}
    where $q^{\mu}$ is an arbitrary four-vector and $\lambda$
    is a scale parameter. In this limit the matrix element
    can be written as
    \begin{equation}\label{eqn:soft_factorisation}
        |\mathcal{M}_{n+1}|^{2} \rightarrow -\dfrac{1}{\lambda^{2}}8\pi \mu^{2\epsilon}\alpha_{s} \sum_{i}\dfrac{1}{p_{i}q}\sum_{k \neq i} \dfrac{p_{k}p_{i}}{(p_{i} + p_{k})q} \; {}_{n}\langle 1, \ldots, n+1 | \bm{T}_{k} \cdot \bm{T}_{i} | 1, \ldots, n+1 \rangle_{n} \, ,
    \end{equation}
    where terms less singular than $1 / \lambda^{2}$ have been
    neglected. The reduced matrix element represented by the
    colour + helicity vector inner product is obtained by removing
    the soft gluon from the $(n+1)$-body matrix element.
    The indices $i$, $j$, and $k$ label partons involved
    in the factorsation process: $i$ is the emitter parton,
    $j$ is the emitted parton, and $k$ is a parton accounting
    for the colour correlations. The scale $\mu$ can be identified as the
    renormalisation scale, and $\bm{T_{i}}$ is the colour-charge
    operator. It acts on the colour vectors in the reduced matrix element
    to give factors of $C_{A}$ and $C_{F}$, hence this factorisation is not exact.

@@

    Similar to (\ref{eqn:soft_factorisation}) where the factorisation was
    only true in the strict soft limit, (\ref{eqn:F_collinear_factorisation})
    and (\ref{eqn:I_collinear_factorisation}) are only
    true in the strict collinear limit. Away from these
    limits care has to be taken to conserve momenta
    via the use of momenta mappings.
